---
title: "Material validation"
subtitle: "Transmission and reflection against DIN/TS 67600"
description: "Photopic and melanopic transmission and reflection coefficients compared with DIN/TS 67600."
date: 2026-10-09
lang: en
engine: knitr
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
    code-fold: show
    code-summary: "Calculation"
    code-overflow: wrap
    embed-resources: true
    html-math-method: mathml
    css: Material_Validation.css
execute:
  echo: true
  message: false
  warning: true
  error: false
  cache: false
params:
  output-dir: ""
---

## Preface

This article extends the [spectral validation](Validation.html) to Spectran's **Transmission / Reflection** module. It compares photopic and melanopic material coefficients with **DIN/TS 67600:2022-08, Tables 6 to 10**. The article shows the calculations, reference values and signed differences.

The scope follows the DIN material tables: transmission under **D65**, and reflection under **EE, A, D65, FL11, LED-B1, LED-B2, LED-B3 and LED-B5**. Each material is compared only under the illuminants for which its DIN table supplies coefficients. Comparisons use documented TUB sample assignments. Materials without an assigned spectrum remain in the reference inventory.

The [TUB dataset, version 2](https://doi.org/10.14279/depositonce-11893.2), is explicitly referenced in DIN/TS 67600, Section 6.2.4.1 and reference [43]. Its [documentation](https://depositonce.tu-berlin.de/bitstreams/fb974995-09e8-4448-a0ee-95ddc66230f6/download), pp. 2, 6 and 9, describes the measurement programme, specimens and transmission data. This source relationship does not provide a complete specimen mapping for DIN Tables 7 and 8.

## Table preparation {#sec-preparation}

### Setup and reference input

All scientific calculations use R and the current Spectran source. Loading the project with `pkgload::load_all()` ensures that the article tests the module in the working copy. Internal module functions are called deliberately here because this article checks their implementation.

```{r}
#| label: setup
#| code-fold: true
#| code-summary: "Setup"

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(ggplot2)
  library(gt)
})

root <- rprojroot::find_root(rprojroot::has_file("DESCRIPTION"))
pkgload::load_all(root, quiet = TRUE)
data_dir <- file.path(root, "vignettes/articles/material-validation-data")
response_labels <- c(vis = "Photopic", mel = "Melanopic")
theme_set(cowplot::theme_cowplot(font_size = 11, font_family = "sans"))

fmt <- function(x, digits = 3) formatC(x, digits = digits, format = "f")
sci_prose <- function(x, digits = 2) {
  if (x == 0) return("0")
  parts <- strsplit(formatC(x, format = "e", digits = digits), "e", fixed = TRUE)[[1]]
  paste0("$", parts[1], "\\times 10^{", as.integer(parts[2]), "}$")
}
```

The [prepared DIN coefficient CSV](material-validation-data/din-67600-material-coefficients.csv) contains one row per material and evaluating illuminant. It records the original German designation, an English translation, transmission or reflection, the photopic coefficient `din_vis`, the melanopic coefficient `din_mel`, and the DIN table and printed page. Both coefficients are dimensionless fractions. The nonvisual quantity considered here is the melanopic coefficient.

The 260 rows cover all 71 materials in Tables 6 to 10. The unglazed reference in Table 6 is not an additional material. CSV preparation is a separate, documented step; rendering this article requires no PDF extraction or table parsing. The [data notes](material-validation-data/data-notes.txt) describe the fields and preparation script.

```{r}
#| label: reference-input

din_file <- file.path(data_dir, "din-67600-material-coefficients.csv")
assignment_file <- file.path(data_dir, "tub-assignments.csv")

din_coefficients <- read.csv(din_file, fileEncoding = "UTF-8") |> as_tibble()
assignments <- read.csv(assignment_file, fileEncoding = "UTF-8") |> as_tibble()

stopifnot(
  nrow(din_coefficients) == 260L,
  n_distinct(din_coefficients$din_id) == 71L,
  !anyDuplicated(din_coefficients[c("din_id", "illuminant")]),
  !anyNA(din_coefficients[c("din_vis", "din_mel")])
)
```

```{r}
#| label: table-style
#| code-fold: true
#| code-summary: "Table and thumbnail styling"

style_table <- function(x) {
  x |>
    opt_align_table_header(align = "left") |>
    tab_options(
      table.font.names = c("Arial", "Helvetica", "sans-serif"),
      table.font.size = px(13), table.width = pct(100),
      data_row.padding = px(6), column_labels.font.weight = "bold",
      table.border.top.width = px(2), table.border.bottom.width = px(2),
      container.overflow.x = "auto"
    )
}

thumbnail <- function(data, alt) {
  plot <- ggplot(data, aes(wavelength, value)) +
    geom_rect(aes(xmin = wavelength - 0.5, xmax = wavelength + 0.5,
      ymin = 0, ymax = value, fill = colour), colour = NA) +
    geom_line(colour = "#222222", linewidth = 0.3) +
    scale_fill_identity() +
    scale_x_continuous(breaks = c(380, 780), expand = expansion(mult = 0)) +
    scale_y_continuous(breaks = c(0, 1), expand = expansion(mult = 0)) +
    coord_cartesian(xlim = c(380, 780), ylim = c(0, 1.03)) +
    labs(x = NULL, y = NULL) +
    theme_classic(base_size = 6) +
    theme(
      axis.line = element_line(linewidth = 0.2),
      axis.ticks = element_line(linewidth = 0.2),
      axis.ticks.length = grid::unit(1, "pt"),
      plot.margin = margin(2, 2, 2, 2)
    )

  image_file <- tempfile(fileext = ".png")
  on.exit(unlink(image_file))
  ggsave(image_file, plot = plot, width = 93 / 96, height = 60 / 96,
    units = "in", dpi = 192, bg = "white", limitsize = FALSE)
  tag <- as.character(gt::local_image(image_file, height = 60))
  sub("<img ", paste0('<img alt="', htmltools::htmlEscape(alt, attribute = TRUE),
    '" '), tag, fixed = TRUE)
}
```

### DIN materials and TUB assignments

The DIN tables contain 12 glazing samples, 11 multi-skin sheets, 21 plastics or filters, and 27 reflecting surfaces. Tables 9 and 10 reuse the same surface samples.

The [assignment CSV](material-validation-data/tub-assignments.csv) links 39 DIN materials to TUB catalogue records. Wood identifiers follow the specimen order in the TUB documentation because the original CSV repeats a wood heading. For S1, DIN describes yellow brick whereas TUB describes beige brick. The comparison tables retain the TUB glazing constructions and this difference in the brick description. Numeric agreement is not used to select specimens.

```{r}
#| label: material-assignments

records <- Spectran:::tub_material_records |>
  select(catalogue_id, display_name_en, source_file, source_column, source_column_header)

materials <- din_coefficients |>
  distinct(din_id, din_name_en, mode, group) |>
  left_join(assignments, by = "din_id", relationship = "one-to-one") |>
  left_join(records, by = "catalogue_id", relationship = "many-to-one") |>
  mutate(
    material = coalesce(sub("^[^·]+ · ", "", display_name_en), din_name_en),
    material = if_else(din_id == 416L, "Yellow brick (TUB: beige)", material),
    curve_available = !is.na(catalogue_id)
  )

stopifnot(nrow(materials) == 71L, !anyNA(materials$material),
  sum(materials$curve_available) == 39L)

coverage <- materials |>
  summarise(Materials = n(), Assigned = sum(curve_available),
    Missing = sum(!curve_available), .by = group)
```

```{r}
#| label: tbl-coverage
#| tbl-cap: "Coverage of the DIN material examples by documented TUB assignments."
#| code-fold: true
#| code-summary: "Format the coverage table"

coverage |>
  gt(rowname_col = "group") |>
  cols_label(Assigned = "Assigned spectra", Missing = "No confirmed spectrum") |>
  fmt_integer(columns = where(is.numeric)) |>
  style_table()
```

### Preparing the DIN illuminants

The source key `CIE:E` represents the equal-energy illuminant labelled **EE** in DIN. Each source is linearly interpolated onto 380–780 nm at 1 nm intervals and scaled to 100 lx using the material module. A scalar change in source intensity cancels in a material coefficient. The thumbnail copies are separately divided by their peak solely to display spectral shape on a common 0–1 scale.

```{r}
#| label: prepare-sources

sources <- din_coefficients |>
  distinct(illuminant) |>
  mutate(
    source_name = if_else(illuminant == "EE", "E", gsub("-", "_", illuminant)),
    source_id = paste0("CIE:", source_name)
  ) |>
  arrange(match(illuminant, c("EE", "A", "D65", "FL11", "LED-B1", "LED-B2", "LED-B3", "LED-B5")))

din_reference <- din_coefficients |>
  select(din_id, illuminant, din_table_ref, page, din_vis, din_mel) |>
  left_join(select(sources, illuminant, source_id),
    by = "illuminant", relationship = "many-to-one") |>
  pivot_longer(c(din_vis, din_mel), names_to = "response",
    names_prefix = "din_", values_to = "din") |>
  select(-illuminant)

source_spectra <- map(sources$source_name, function(name) {
  raw <- Spectran::examplespectra$CIE |>
    filter(between(Wellenlaenge, 380, 780)) |>
    select(Wellenlaenge, all_of(name))
  stopifnot(all(is.finite(as.matrix(raw))), !anyDuplicated(raw[[1]]),
    min(raw[[1]]) == 380, max(raw[[1]]) == 780)

  spectrum <- tibble(
    Wellenlaenge = 380:780,
    Bestrahlungsstaerke = approx(raw[[1]], raw[[2]], xout = 380:780)$y
  )
  Spectran:::material_scale_light(spectrum, 100)
}) |>
  setNames(sources$source_id)

stopifnot(all(map_lgl(source_spectra, function(s) {
  all(is.finite(s$Bestrahlungsstaerke)) && all(s$Bestrahlungsstaerke >= 0)
})))
```

## Calculation {#sec-calculation}

For incident spectrum $S(\lambda)$, material spectrum $c(\lambda)$ and weighting function $w(\lambda)$, the coefficient is

$$
c_w(S)=\frac{\int_{380}^{780} S(\lambda)c(\lambda)w(\lambda)\,\mathrm d\lambda}
{\int_{380}^{780} S(\lambda)w(\lambda)\,\mathrm d\lambda}.
$$ {#eq-coefficient}

Here $c=\tau$ for transmission and $c=\rho$ for reflection. The photopic weighting is $V(\lambda)$; the melanopic weighting is $s_{\mathrm{mel}}(\lambda)$. Both are taken from `Specs$AS_wide`. For illuminants other than D65, DIN also uses the designation $k_{\mathrm{mel}}$ for the melanopic material factor. We compare the weighted material coefficients, not MDER or effective MDER.

Spectran linearly interpolates the TUB material samples from 5 nm to 1 nm and sums over 380–780 nm, including both endpoints. The common wavelength increment cancels in the quotient. The outgoing spectrum is not renormalised when calculating a material coefficient.

```{r}
#| label: prepare-materials

mapped <- filter(materials, curve_available)
material_preparation <- map(mapped$catalogue_id, function(id) {
  curve <- Spectran:::transmission_catalogue_record(id)$curve
  stopifnot(identical(as.numeric(curve$wavelength_nm), seq(380, 780, 5)))

  raw <- data.frame(wavelength_nm = curve$wavelength_nm, value = curve$transmittance)
  prepared <- Spectran:::prepare_transmission_curve(raw, scale = "fraction")$completed
  stopifnot(all(is.finite(prepared$transmittance)),
    all(prepared$status %in% c("supplied", "interpolated")))
  prepared
}) |>
  setNames(mapped$catalogue_id)

weights <- list(
  vis = Spectran::Specs$AS_wide[["V(lambda)"]],
  mel = Spectran::Specs$AS_wide$Melanopsin
)
stopifnot(identical(as.numeric(Spectran::Specs$AS_wide$Wellenlaenge), as.numeric(380:780)))
```

### Worked example: DIN 106 / TUB G6 under D65

The numerator describes the transmitted, weighted spectrum. The denominator describes the incident, weighted spectrum. Their ratio is dimensionless.

```{r}
#| label: worked-example

s <- source_spectra[["CIE:D65"]]$Bestrahlungsstaerke
c_material <- material_preparation[["tub:transmission:G6"]]$transmittance

example_terms <- imap(weights, function(w, response) {
  numerator <- sum(s * c_material * w)
  denominator <- sum(s * w)
  tibble(
    Response = unname(response_labels[response]),
    Numerator = numerator, Denominator = denominator,
    Coefficient = numerator / denominator
  )
}) |>
  list_rbind()
```

```{r}
#| label: tbl-example
#| tbl-cap: "Explicit weighted sums for DIN 106 / TUB G6 under D65."
#| code-fold: true
#| code-summary: "Format the worked example"

example_terms |>
  gt(rowname_col = "Response") |>
  fmt_number(columns = where(is.numeric), decimals = 6) |>
  style_table()
```

The resulting photopic coefficient is `r fmt(example_terms$Coefficient[1], 6)` and the melanopic coefficient is `r fmt(example_terms$Coefficient[2], 6)`.

### Checking the material module

For every confirmed material–illuminant pair in the DIN tables, we execute both the module and the explicit weighted sums. The module field `active_metrics$comparison_value` supplies the outgoing-to-incident ratio. This is an implementation check, not an independent test of the input spectra.

```{r}
#| label: calculate-confirmed

coefficient_values <- function(catalogue_id, source_id, mode) {
  spectrum <- source_spectra[[source_id]]
  curve <- material_preparation[[catalogue_id]]
  result <- Spectran:::calculate_material_result(spectrum, curve, tolower(mode))
  metric_ids <- c(vis = "photopic_illuminance", mel = "melanopic_irradiance")

  imap(weights, function(w, response) {
    numerator <- sum(spectrum$Bestrahlungsstaerke * curve$transmittance * w)
    denominator <- sum(spectrum$Bestrahlungsstaerke * w)
    value <- result$active_metrics$comparison_value[
      result$active_metrics$metric_id == metric_ids[[response]]
    ]
    stopifnot(length(value) == 1L, is.finite(value), denominator > 0)

    tibble(
      catalogue_id, source_id, response,
      spectran = value, numerator, denominator,
      explicit = numerator / denominator
    )
  }) |>
    list_rbind()
}

pairs <- din_reference |>
  distinct(din_id, source_id) |>
  inner_join(select(mapped, din_id, catalogue_id, mode),
    by = "din_id", relationship = "many-to-one")

calculated <- pmap(pairs, function(din_id, source_id, catalogue_id, mode) {
  coefficient_values(catalogue_id, source_id, mode) |>
    mutate(din_id = din_id)
}) |>
  list_rbind()

internal_max <- max(abs(calculated$spectran - calculated$explicit))
stopifnot(internal_max <= 1e-12)
```

The module and the explicit sums agree within $10^{-12}$ for all `r nrow(calculated)` coefficients (`r nrow(pairs)` material–illuminant pairs). The maximum absolute difference is `r sci_prose(internal_max)`.

### Differences and rounding

We report $\Delta=c_{\mathrm{Spectran}}-c_{\mathrm{DIN}}$, its percentage-point form $100\Delta$, and the relative difference $100\Delta/c_{\mathrm{DIN}}$. Positive values mean that Spectran is higher. Relative differences are undefined for a zero DIN value.

The strict rounding interval is **±0.0005** for a DIN coefficient printed to three decimals. A second, project-defined **reproduction limit of ±0.001** (±0.10 percentage points) also allows for TUB spectral coefficients rounded to three decimals. With nonnegative weights, rounding each input coefficient to the nearest 0.001 can change the weighted mean by at most 0.0005. Adding the two rounding bounds gives 0.001. This is neither a DIN conformity tolerance nor measurement uncertainty, and it does not explain the cause of any individual difference.

```{r}
#| label: compare-confirmed

material_details <- materials |>
  select(din_id, material, mode, group, catalogue_id, curve_available)
calculation_details <- calculated |>
  select(din_id, source_id, response, spectran, numerator, denominator, explicit)

comparison <- din_reference |>
  left_join(material_details, by = "din_id", relationship = "many-to-one") |>
  left_join(calculation_details,
    by = c("din_id", "source_id", "response"), relationship = "one-to-one") |>
  left_join(select(sources, source_id, illuminant),
    by = "source_id", relationship = "many-to-one") |>
  mutate(
    difference = spectran - din,
    difference_pp = 100 * difference,
    relative_difference_percent = if_else(din != 0, 100 * difference / din, NA_real_),
    within_rounding = abs(difference) <= 0.0005,
    within_budget = abs(difference) <= 0.001,
    status = case_when(
      !curve_available ~ "No confirmed spectrum",
      within_rounding ~ "Within rounding",
      within_budget ~ "Within reproduction limit",
      TRUE ~ "Outside reproduction limit"
    )
  )

stopifnot(nrow(comparison) == 520L)
paired <- filter(comparison, curve_available)
wide <- comparison |>
  select(din_id, material, mode, catalogue_id, curve_available, source_id,
    illuminant, response, din, spectran, difference_pp, relative_difference_percent) |>
  pivot_wider(
    names_from = response,
    values_from = c(din, spectran, difference_pp, relative_difference_percent)
  )
```

## Results {#sec-results}

The overview includes only confirmed assignments with both DIN and Spectran values. The result tables below pair each material curve with the applicable illuminant curve; the material plots retain absolute coefficients on a 0–1 scale, whereas illuminant plots show peak-normalised spectral power. All thumbnails span 380–780 nm. Their filled areas use Spectran's `Lang` wavelength palette. These colours indicate wavelength and do not predict the perceived colour of a material or reproduce monochromatic light faithfully on a screen.

```{r}
#| label: result-summary

summary_by_response <- paired |>
  summarise(
    Coefficients = n(), Rounding = sum(within_rounding),
    Reproduced = sum(within_budget), Outside = sum(!within_budget),
    median_abs_pp = median(abs(difference_pp)),
    max_abs_pp = max(abs(difference_pp)),
    .by = c(mode, response)
  )

summary_by_source <- paired |>
  summarise(
    Coefficients = n(), Outside = sum(!within_budget),
    max_abs_pp = max(abs(difference_pp)),
    .by = c(mode, illuminant, response)
  )
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Agreement for confirmed assignments. The reproduction count includes the stricter rounding count."
#| code-fold: true
#| code-summary: "Format the overview"

summary_by_response |>
  mutate(response = unname(response_labels[response])) |>
  gt(groupname_col = "mode", rowname_col = "response") |>
  cols_label(Coefficients = "n", Rounding = html("Within<br>±0.0005"),
    Reproduced = html("Within<br>±0.001"), Outside = "Outside",
    median_abs_pp = html("Median |Δ|<br>(pp)"), max_abs_pp = html("Max. |Δ|<br>(pp)")) |>
  fmt_integer(columns = c(Coefficients, Rounding, Reproduced, Outside)) |>
  fmt_number(columns = c(median_abs_pp, max_abs_pp), decimals = 4) |>
  style_table() |>
  tab_source_note("pp = percentage points.")
```

```{r}
#| label: fig-overview
#| fig-cap: "Signed differences for confirmed assignments. Dashed lines mark the project reproduction limit of ±0.10 percentage points; the two panels use the same vertical scale."
#| fig-alt: "Two panels show photopic and melanopic Spectran-minus-DIN differences by illuminant. Circles identify transmission and triangles reflection. Horizontal dashed lines mark plus and minus 0.10 percentage points."
#| fig-width: 7.2
#| fig-height: 3.6

figure_overview_data <- paired |>
  mutate(
    illuminant = factor(illuminant, levels = sources$illuminant),
    response = factor(response, levels = names(response_labels), labels = response_labels)
  )

ggplot(figure_overview_data, aes(illuminant, difference_pp, colour = mode, shape = mode)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_hline(yintercept = c(-0.1, 0.1), linetype = "dashed",
    colour = "grey45", linewidth = 0.35) +
  geom_point(size = 1.7, alpha = 0.65, position = position_dodge(width = 0.3)) +
  facet_wrap(~ response, nrow = 1) +
  scale_colour_manual(values = c(Reflection = "#264E86", Transmission = "#B04A32")) +
  scale_shape_manual(values = c(Reflection = 17, Transmission = 16)) +
  labs(x = "Illuminant", y = "Spectran − DIN (percentage points)", colour = NULL, shape = NULL) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: result-table-format
#| code-fold: true
#| code-summary: "Prepare the two spectral thumbnails and format the result tables"

wavelength_colours <- tibble(
  wavelength = 380:780,
  colour = unname(Spectran::ColorP$Lang[as.character(380:780)])
)
stopifnot(!anyNA(wavelength_colours))

source_thumbnails <- map(source_spectra, function(s) {
  tibble(wavelength = s$Wellenlaenge,
    value = s$Bestrahlungsstaerke / max(s$Bestrahlungsstaerke)) |>
    left_join(wavelength_colours, by = "wavelength", relationship = "one-to-one")
})
material_thumbnails <- map(material_preparation, function(curve) {
  tibble(wavelength = curve$wavelength_nm, value = curve$transmittance) |>
    left_join(wavelength_colours, by = "wavelength", relationship = "one-to-one")
})

source_images <- imap_chr(source_thumbnails, function(data, id) {
  label <- sources$illuminant[match(id, sources$source_id)]
  image <- thumbnail(data, paste(label, "relative spectral power, 380–780 nm"))
  paste0(image, '<br><span class="spectrum-label">', label, '</span>')
})
material_images <- imap_chr(material_thumbnails, function(data, id) {
  label <- materials$material[match(id, materials$catalogue_id)]
  thumbnail(data, paste(label, "material coefficient, 380–780 nm"))
})

signed_cell <- function(x, digits = 3, mark_limit = FALSE) {
  map_chr(x, function(value) {
    if (is.na(value)) return("NA")
    # Suppress a negative sign only when the displayed value rounds to zero.
    displayed <- if (round(value, digits) == 0) 0 else value
    label <- formatC(displayed, format = "f", digits = digits, flag = "+")
    if (mark_limit && abs(value) > 0.1) {
      paste0('<span class="outside-limit"><strong>', label, '</strong></span>')
    } else {
      label
    }
  })
}

result_table <- function(data) {
  table_data <- data |>
    mutate(
      identity = paste0(din_id, "<br><span class='sample-id'>",
        sub("tub:(transmission|reflection):", "", catalogue_id), "</span>"),
      material_plot = unname(material_images[catalogue_id]),
      source_plot = unname(source_images[source_id]),
      response = "vis<br>mel"
    ) |>
    select(identity, material, material_plot, source_plot, response,
      din_vis, din_mel, spectran_vis, spectran_mel, difference_pp_vis,
      difference_pp_mel, relative_difference_percent_vis, relative_difference_percent_mel)

  table_data |>
    gt() |>
    fmt_number(columns = c(din_vis, din_mel), decimals = 3) |>
    fmt_number(columns = c(spectran_vis, spectran_mel), decimals = 5) |>
    gt::fmt(columns = c(difference_pp_vis, difference_pp_mel),
      fns = function(x) signed_cell(x, mark_limit = TRUE)) |>
    gt::fmt(columns = c(relative_difference_percent_vis, relative_difference_percent_mel),
      fns = signed_cell) |>
    sub_missing(missing_text = "NA") |>
    cols_merge(columns = c(din_vis, din_mel), pattern = "{1}<br>{2}") |>
    cols_merge(columns = c(spectran_vis, spectran_mel), pattern = "{1}<br>{2}") |>
    cols_merge(columns = c(difference_pp_vis, difference_pp_mel), pattern = "{1}<br>{2}") |>
    cols_merge(columns = c(relative_difference_percent_vis, relative_difference_percent_mel),
      pattern = "{1}<br>{2}") |>
    fmt_markdown(columns = c(identity, material_plot, source_plot, response)) |>
    cols_label(
      identity = html("DIN<br>TUB"), material = "Material description",
      material_plot = "Material", source_plot = "Illuminant", response = "Weight",
      din_vis = "DIN", spectran_vis = "Spectran",
      difference_pp_vis = html("Δ<br>(pp)"),
      relative_difference_percent_vis = html("Δ<br>(%)")
    ) |>
    cols_align(align = "left", columns = material) |>
    cols_align(align = "center", columns = c(identity, material_plot, source_plot, response)) |>
    cols_align(align = "right",
      columns = c(din_vis, spectran_vis, difference_pp_vis, relative_difference_percent_vis)) |>
    cols_width(
      identity ~ px(42), material ~ px(154), material_plot ~ px(80), source_plot ~ px(80),
      response ~ px(45), din_vis ~ px(62), spectran_vis ~ px(69),
      difference_pp_vis ~ px(62), relative_difference_percent_vis ~ px(63)
    ) |>
    style_table() |>
    # Preserve paired cell formatting and column widths through Quarto/pkgdown.
    tab_options(table.font.size = px(12), data_row.padding = px(4),
      quarto.disable_processing = TRUE) |>
    tab_source_note(html(paste0(
      "Within each cell: <b>vis</b> (photopic), then <b>mel</b> (melanopic). ",
      "Δ = Spectran − DIN. pp = percentage points; % = relative difference. ",
      "Bold red Δ values exceed ±0.10 pp. ",
      "Thumbnails: 380–780 nm; material 0–1; illuminant peak = 1."
    )))
}
```

### Transmission {#sec-transmission}

#### Confirmed glazing assignments

Table 6 supplies D65 coefficients for the 12 glazing samples. These are the confirmed transmission comparisons. Five decimal places are retained for Spectran to make small differences visible; they do not imply equivalent measurement accuracy.

::: {.material-table-key}
Each numeric cell lists **photopic (vis)** above **melanopic (mel)** values. Δ = Spectran − DIN; **pp** = percentage points; **%** = relative difference. Bold red Δ values exceed ±0.10 pp.

<span class="material-scroll-hint">Scroll the table sideways to compare all columns. The compact DIN/TUB identifiers stay visible.</span>
:::

```{r}
#| label: tbl-transmission
#| tbl-cap: "Transmission under D65: DIN Table 6 and the assigned TUB glazing spectra."
#| code-fold: true
#| code-summary: "R code: transmission table"

wide |>
  filter(mode == "Transmission", curve_available) |>
  arrange(din_id) |>
  result_table()
```

#### Materials without an assigned spectrum

No confirmed TUB spectrum is assigned to the following 32 DIN materials. A missing assignment can mean that the named material is absent from the source or that the available descriptions do not establish specimen identity. All entries below are evaluated under D65 in DIN; no substitute spectra are used.

```{r}
#| label: missing-inventory

missing_inventory <- wide |>
  filter(!curve_available) |>
  left_join(select(materials, din_id, group), by = "din_id", relationship = "one-to-one") |>
  arrange(din_id) |>
  select(group, din_id, material, din_vis, din_mel)

stopifnot(nrow(missing_inventory) == 32L)
```

<p class="material-scroll-hint material-table-key">Scroll the table sideways to compare both coefficients. The DIN IDs stay visible.</p>

```{r}
#| label: tbl-missing
#| tbl-cap: "DIN materials without an assigned TUB spectrum, with their D65 reference coefficients. Multi-skin sheets are from Table 7; plastics and filters are from Table 8."
#| code-fold: true
#| code-summary: "Format the missing-material inventory"

missing_inventory |>
  gt(groupname_col = "group") |>
  cols_label(din_id = "DIN ID", material = "Material",
    din_vis = "DIN photopic", din_mel = "DIN melanopic") |>
  fmt_number(columns = c(din_vis, din_mel), decimals = 3) |>
  style_table()
```

### Reflection {#sec-reflection}

Tables 9 and 10 provide eight illuminants for the same 27 reflecting surfaces. The tabs follow the DIN order. Each table reports both weighting functions and retains all samples, including the FL11 differences.

::: {.material-table-key}
Each numeric cell lists **photopic (vis)** above **melanopic (mel)** values. Δ = Spectran − DIN; **pp** = percentage points; **%** = relative difference. Bold red Δ values exceed ±0.10 pp.

<span class="material-scroll-hint">Scroll the table sideways to compare all columns. The compact DIN/TUB identifiers stay visible.</span>
:::

::: {.material-tabs}

<ul class="nav nav-tabs" role="tablist" aria-label="Reflection illuminants">
<li class="nav-item" role="presentation"><button class="nav-link active" id="reflection-ee-tab" data-bs-toggle="tab" data-bs-target="#reflection-ee-pane" type="button" role="tab" aria-controls="reflection-ee-pane" aria-selected="true">EE</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-a-tab" data-bs-toggle="tab" data-bs-target="#reflection-a-pane" type="button" role="tab" aria-controls="reflection-a-pane" aria-selected="false">A</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-d65-tab" data-bs-toggle="tab" data-bs-target="#reflection-d65-pane" type="button" role="tab" aria-controls="reflection-d65-pane" aria-selected="false">D65</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-fl11-tab" data-bs-toggle="tab" data-bs-target="#reflection-fl11-pane" type="button" role="tab" aria-controls="reflection-fl11-pane" aria-selected="false">FL11</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-led-b1-tab" data-bs-toggle="tab" data-bs-target="#reflection-led-b1-pane" type="button" role="tab" aria-controls="reflection-led-b1-pane" aria-selected="false">LED-B1</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-led-b2-tab" data-bs-toggle="tab" data-bs-target="#reflection-led-b2-pane" type="button" role="tab" aria-controls="reflection-led-b2-pane" aria-selected="false">LED-B2</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-led-b3-tab" data-bs-toggle="tab" data-bs-target="#reflection-led-b3-pane" type="button" role="tab" aria-controls="reflection-led-b3-pane" aria-selected="false">LED-B3</button></li>
<li class="nav-item" role="presentation"><button class="nav-link" id="reflection-led-b5-tab" data-bs-toggle="tab" data-bs-target="#reflection-led-b5-pane" type="button" role="tab" aria-controls="reflection-led-b5-pane" aria-selected="false">LED-B5</button></li>
</ul>

::: {.tab-content}

::: {#reflection-ee-pane .tab-pane .fade .show .active role="tabpanel" aria-labelledby="reflection-ee-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-ee
#| tbl-cap: "Reflection under EE, DIN Table 9."
#| code-fold: true
#| code-summary: "R code: EE reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:E") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-a-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-a-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-a
#| tbl-cap: "Reflection under A, DIN Table 9."
#| code-fold: true
#| code-summary: "R code: A reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:A") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-d65-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-d65-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-d65
#| tbl-cap: "Reflection under D65, DIN Table 9."
#| code-fold: true
#| code-summary: "R code: D65 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:D65") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-fl11-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-fl11-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-fl11
#| tbl-cap: "Reflection under FL11, DIN Table 9. Bold red differences exceed the reproduction limit."
#| code-fold: true
#| code-summary: "R code: FL11 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:FL11") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-led-b1-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-led-b1-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-led-b1
#| tbl-cap: "Reflection under LED-B1, DIN Table 10."
#| code-fold: true
#| code-summary: "R code: LED-B1 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:LED_B1") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-led-b2-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-led-b2-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-led-b2
#| tbl-cap: "Reflection under LED-B2, DIN Table 10."
#| code-fold: true
#| code-summary: "R code: LED-B2 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:LED_B2") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-led-b3-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-led-b3-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-led-b3
#| tbl-cap: "Reflection under LED-B3, DIN Table 10."
#| code-fold: true
#| code-summary: "R code: LED-B3 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:LED_B3") |>
  arrange(din_id) |>
  result_table()
```

:::

::: {#reflection-led-b5-pane .tab-pane .fade role="tabpanel" aria-labelledby="reflection-led-b5-tab" tabindex="0"}

```{r}
#| label: tbl-reflection-led-b5
#| tbl-cap: "Reflection under LED-B5, DIN Table 10."
#| code-fold: true
#| code-summary: "R code: LED-B5 reflection table"
wide |>
  filter(mode == "Reflection", source_id == "CIE:LED_B5") |>
  arrange(din_id) |>
  result_table()
```

:::

:::

:::

## Conclusion {#sec-conclusion}

```{r}
#| label: computed-conclusion
#| output: asis
#| code-fold: true
#| code-summary: "Derive the conclusion from this run"

outside <- filter(paired, !within_budget)
cat(sprintf(paste0(
  "**%d of %d confirmed coefficients** lie within the project reproduction limit of ±0.001; ",
  "%d already lie within the stricter rounding interval of ±0.0005. ",
  "The remaining %d coefficients exceed the reproduction limit.\n\n"
), sum(paired$within_budget), nrow(paired), sum(paired$within_rounding), nrow(outside)))

if (nrow(outside)) {
  affected <- outside |> distinct(mode, response, illuminant)
  if (nrow(affected) == 1L) {
    cat(sprintf("All exceedances concern **%s %s under %s**.\n\n",
      tolower(response_labels[affected$response]), tolower(affected$mode), affected$illuminant))
  }

  worst <- slice_max(outside, abs(difference_pp), n = 1, with_ties = FALSE)
  cat(sprintf(paste0(
    "The largest absolute difference is **DIN %d, %s, %s**: ",
    "DIN %s, Spectran %s, Δ %s percentage points.\n\n"
  ), worst$din_id, worst$material, worst$illuminant,
    fmt(worst$din, 3), fmt(worst$spectran, 6), fmt(worst$difference_pp, 4)))
}
```

The module and the explicit weighted sums agree for all tested inputs.

The DIN photopic reflection coefficients for FL11 have a known deviation from the correct values. According to the author of the DIN source, Spectran calculates these coefficients correctly, and a correction of the DIN values is planned for the next edition of DIN/TS 67600.[^lang-2026]

[^lang-2026]: Private correspondence between Johannes Zauner and Dieter Lang, 8 October 2026.

For the remaining **32 DIN materials**, no confirmed spectrum is assigned. Two integrated coefficients are insufficient to reconstruct a unique material spectrum. A complete spectral validation of all 71 DIN materials is therefore not supported by the available specimen documentation.

These checks assess numerical reproduction for documented spectra and calculation conventions. They do not replace independent material measurements. Applying the values to an actual room or product still requires the relevant specimen properties and measurement geometry.

## Reproduction {#sec-reproduce}

### Inputs and exports

This article runs from the Spectran source tree. Supplying `output-dir`, or the environment variable `SPECTRAN_MATERIAL_OUTPUT_DIR`, writes the unrounded results as CSV and an R object. `comparison-long.csv` includes weighted numerators and denominators; `comparison-wide.csv` places the two responses side by side. `material-map.csv` records the source column for every confirmed assignment. The data behind the overview figure and both sets of thumbnails are also exported as CSV. Without an explicit output directory, these files are written to `data-raw/intermediate/material-validation/`.

```{r}
#| label: exports
#| code-fold: true
#| code-summary: "Export results and record input checksums"

input_files <- file.path(root, c(
  "DESCRIPTION", "renv.lock", "R/sysdata.rda", "data/Specs.rda", "data/ColorP.rda",
  "data/examplespectra.rda", "data/examplespectra_descriptor.rda",
  "R/spectral_metrics.R", "R/transmission_core.R", "R/material_core.R",
  "R/material_light_level.R", "R/transmission_catalogue.R", "R/import_verifier.R",
  "inst/extdata/tub/Spectral_transmittance.csv", "inst/extdata/tub/Spectral_reflectance.csv",
  "vignettes/articles/Material_Validation.qmd", "vignettes/articles/Material_Validation.css"
))
input_files <- c(input_files, din_file, assignment_file)
stopifnot(all(file.exists(input_files)))
input_hashes <- tibble(
  File = substring(input_files, nchar(root) + 2L),
  MD5 = unname(tools::md5sum(input_files))
)

output_dir <- params[["output-dir"]]
if (!nzchar(output_dir)) output_dir <- Sys.getenv("SPECTRAN_MATERIAL_OUTPUT_DIR", "")
if (!nzchar(output_dir)) output_dir <- file.path(root, "data-raw/intermediate/material-validation")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

material_thumbnail_data <- list_rbind(material_thumbnails, names_to = "catalogue_id")
source_thumbnail_data <- list_rbind(source_thumbnails, names_to = "source_id")
outputs <- list(
  "comparison-long" = comparison, "comparison-wide" = wide,
  "summary-by-response" = summary_by_response, "summary-by-source" = summary_by_source,
  "material-map" = materials, "source-inventory" = sources,
  "missing-materials" = missing_inventory, "input-hashes" = input_hashes,
  "figure-overview-data" = figure_overview_data,
  "material-thumbnails-data" = material_thumbnail_data,
  "source-thumbnails-data" = source_thumbnail_data
)
iwalk(outputs, function(data, name) {
  write.csv(data, file.path(output_dir, paste0(name, ".csv")), row.names = FALSE, na = "")
})

saveRDS(list(calculated = calculated, comparison = comparison,
  example_terms = example_terms, inputs = input_hashes),
  file.path(output_dir, "validation-results.rds"))
writeLines(capture.output(sessionInfo()), file.path(output_dir, "sessionInfo.txt"))
```

<details>
<summary>Input files and checksums for this render</summary>

```{r}
#| label: tbl-input-hashes
#| tbl-cap: "Input provenance for this render. Paths are relative to the package root where possible."
#| code-fold: true
#| code-summary: "Format the input manifest"

input_hashes |>
  gt() |>
  style_table() |>
  tab_options(table.font.size = px(11))
```

</details>

### Render commands

The prepared CSV is included with the article. From the package root, render the standalone article with:

```bash
quarto render vignettes/articles/Material_Validation.qmd
```

To choose a directory for the unrounded results and plot data:

```bash
quarto render vignettes/articles/Material_Validation.qmd \
  -P output-dir:/path/material-validation
```

The same Quarto source is registered in pkgdown's Articles menu. The website build requires **pkgdown ≥ 2.2.1**, the R package **quarto**, and a Quarto CLI:

```r
pkgdown::build_site(preview = FALSE)
```

CSV preparation can be repeated separately using `data-raw/prepare_din_material_coefficients.R` and a local, licensed text extraction. Neither the licensed PDF nor its extracted text is needed to render the article.

### Software

The version table describes the R environment and Quarto tools used for this render.

```{r}
#| label: software-versions
#| code-fold: true
#| code-summary: "Collect the software versions"

packages <- c("Spectran", "pkgload", "dplyr", "tidyr", "purrr", "ggplot2", "gt",
  "knitr", "rmarkdown", "pkgdown", "quarto")
software_versions <- tibble(
  Component = c("R", "Quarto CLI", packages),
  Version = c(as.character(getRversion()), system2("quarto", "--version", stdout = TRUE),
    map_chr(packages, function(p) as.character(packageVersion(p))))
)
write.csv(software_versions, file.path(output_dir, "software-versions.csv"), row.names = FALSE)
```

```{r}
#| label: tbl-software
#| tbl-cap: "Software versions used to render this article."
#| code-fold: true
#| code-summary: "Format the software record"

software_versions |>
  gt() |>
  style_table()
```

The reference basis is DIN/TS 67600:2022-08, Section 6.2.4 and Tables 6–10, and the [TUB dataset, version 2](https://doi.org/10.14279/depositonce-11893.2). The weighting functions and bundled illuminants are also described in [Underlying values](Values.html).
