---
title: "Validation"
date: 2023-11-03
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  code_folding = "hide",
  comment = "#>"
)
```

## Preface

This document compares the results for ten spectra across the `Spectral Analysis` Shiny application, the online app [`luox`](https://luox.app) by Manuel Spitschan, and the free [`CIE S 026 alpha-opic Toolbox`](https://files.cie.co.at/CIE%20S%20026%20alpha-opic%20Toolbox%20User%20Guide.pdf). The main goal of this document is to validate the `Spectral Analysis` application against the methods the CIE has provided or already validated. We will look at results for `illuminance`, `α-opic irradiance`, and `α-opic equivalent daylight (D65) illuminance`. We will also compare the `irradiance` (only with the toolbox) and `Correlated Color Temperature (CCT)` and `Color-Rendering Index (CRI)` (only with the luox app), which are only available in one of the two validated sources. All three sources offer more parameters, but those are either not part of both the shiny application and a validated source or are derivatives of the above-mentioned parameters. Some spectral data files had negative input values (coming straight from the spectrometer export after measurement). The `Shiny` app replaces these values with zero, whereas the `CIE toolbox` gives an error for these values. Here, the values were manually set to zero in the toolbox user input mask. The `luox` app gives no error for negative values, but it is not exactly known how the app deals with those values (see below under @sec-conclusion for more on that).

The results for the `CIE Toolbox` and `luox` were taken on `21 October 2022` with their respective current version at that date.

## Table Preparation

The following code chunks prepare the tables shown below. The first chunk loads the necessary libraries:

<details><summary> Setup </summary> 
```{r, echo = TRUE}

#Setup code, data import, initial data selection to get to one comparison file
#read all libraries
library(tidyverse)
library(readxl)
library(magick)
library(gt)
library(here)
library(cowplot)

#Plottheme
theme_set(theme_cowplot(font_size = 10, font_family = "sans"))

```
</details> 

The following chunk loads all data:

<details><summary> Data Import </summary> 
```{r}
#Data Import -------------------

##read all filenames and paths of spectra
spectra <- 
  tibble(
    file_names = list.files("./Original_Spectra"),
    file_path = paste0("./Original_Spectra/", list.files("./Original_Spectra"))
    )

##read the csv-files to that spectra
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(Spectrum = list(read_csv(file_path)))

##create a name column 
spectra <- 
  spectra %>% 
  mutate(spectrum_name = str_replace(file_names, ".csv", ""))

##read the results from the luox app
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(
    luox = list(
      read_csv(
        paste0("./Results_Luox_2022-10-21/", spectrum_name, 
               "/download-calc.csv")
        )
      )
    )

##read the results from the shiny app
###filepath
excel_file_path <- function(spectrum_name) {
  paste0(
    "./Results_ShinyApp/",
    {{ spectrum_name }}, 
    "/",
    {{ spectrum_name }}, 
    "_9_2022-10-20.xlsx"
    )
}
###list for each worksheet in the excel-file
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(
    shiny = list(
      list(
        Radiometrie = read_xlsx(
          excel_file_path(spectrum_name), 
          sheet = "Radiometrie"
          ),
        Photometrie = read_xlsx(
          excel_file_path(spectrum_name), 
          sheet = "Photometrie"
          ),
        Alpha = read_xlsx(
          excel_file_path(spectrum_name), 
          sheet = "Alpha-opisch"
          )
        )
      )
    ) 

##read the results from the CIE toolbox
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(
    toolbox = list(
      read_xlsx(
        paste0(
          "./Results_CIE_Toolbox/", 
          spectrum_name, 
          "/CIE S 026 alpha-opic Toolbox.xlsx"
          ), 
        sheet = "Outputs"
        )
      )
    )


```
</details> 

The following chunk takes the relevant data for comparison out from the sources and puts them into one table per comparison spectrum:

<details><summary> Initial Data wrangling </summary> 
```{r}

#Initial Data wrangling -------------------
##take the relevant datapoints from the luox results
### the relevant data in the luox results are in column 2, rows 1, 6 to 15, 22, 
### and 24
locations_luox <- c(1, 6:15, 22, 24)
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(
    excerpt = list(
      tibble(
        Name = luox %>% pull(1) %>% .[locations_luox],
        Results_luox = luox %>% pull(2) %>% .[locations_luox]
        )
      )
    )

###Add a row for irradiance, which is missing in the luox output
spectra <- 
  spectra %>% 
  mutate(
    excerpt = list(
      rbind(
        excerpt[1:11,], 
        c("Irradiance (mW ⋅ m⁻²)", NA), 
        excerpt[12:13,]
        )
      ),
    excerpt = list(
      excerpt %>% mutate(Results_luox = as.numeric(Results_luox))
    )
    )

##extract the relevant datapoints from the shiny app
### the relevant data in the shiny app results are in the list 
### - "Photometrie", column 3, row 1, 
### - "Alpha", column 3 to 7, row 1 to 2 (the order has to be adjusted in order 
###to match the luox data frame)
### - "Radiometrie, column 3, row 1, 
### - "Photometrie", column 3, row 4 to 5
shiny_extract <- function(data, sheet, column = Wert, rows) {
  data %>% .[[sheet]] %>% pull({{ column }}) %>% .[rows]
}
spectra <- 
  spectra %>% 
  rowwise() %>% 
  mutate(
    excerpt = list(
      cbind(
        excerpt[1],
        tibble(
        Results_shiny = as.numeric(
          c(
            shiny_extract(
              data = shiny, sheet = "Photometrie", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Cyanopsin", rows = 2),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Chloropsin", rows = 2),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Erythropsin", rows = 2),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Rhodopsin", rows = 2),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Melanopsin", rows = 2),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Cyanopsin", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Chloropsin", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Erythropsin", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Rhodopsin", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Alpha", column = "Melanopsin", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Radiometrie", rows = 1),
            shiny_extract(
              data = shiny, sheet = "Photometrie", rows = c(4, 5))
            )
          )
        ),
        excerpt[2]
        )
      )
  )

##extract the relevant datapoints from the CIE Toolbox
### the relevant data in the toolbox are in 
### column 3, row 14
### column 1 to 5, row 20-> needs to be multiplied by a factor of 1000 to be in 
### mW, to which the other sources are scaled.
### column 1 to 5, row 32
### column 1, row 14 -> needs to be multiplied by a factor of 1000 to be in mW, 
### to which the other sources are scaled.
spectra <- 
  spectra %>% 
  mutate(
    excerpt = list(
      cbind(
        excerpt,
        tibble(
          Results_toolbox = as.numeric(
            c(
              toolbox %>% pull(3) %>% .[14],
              toolbox %>% {as.vector(.[20,])} %>% as.numeric %>% 
                magrittr::multiply_by(1000),
              toolbox %>% {as.vector(.[32,])},
              toolbox %>% pull(1) %>% .[14] %>% as.numeric %>% 
                magrittr::multiply_by(1000),
              NA, NA
              )
            )
          )
        )
      )
    )






```
</details> 

The following chunk transforms the comparison tables for all spectra into one comprehensive table.

<details><summary> Putting the table together </summary> 
```{r}
#Putting the table together -------------------
##create a function that calculates the relative difference between the shiny 
##app-results, and another source
Deviation <- function(Results, Results2){
  if(!is.na({{ Results }}) & !is.na({{ Results2 }})){
  res <- 1- {{ Results }} / {{ Results2 }}
  res2 <- vec_fmt_scientific(res)
  if(res < 0) {
    paste0('<div style="color:red">', res2, '</div>')
  }
  else if(res == 0 ) {
    paste0('<div style="color:green">', res2, '</div>')
  }
  else {
    paste0('<div style="color:blue">', res2, '</div>')
  }

  }
  else NA
}

##new dataframe, unnested data, columns for relative difference
Results <- 
  spectra %>% 
  select(spectrum_name, excerpt) %>% 
  unnest(excerpt) %>% 
  rowwise() %>% 
  mutate(Dev_luox = Deviation(Results_luox, Results_shiny),
         Dev_toolbox = Deviation(Results_toolbox, Results_shiny)
         )

##pivoting the dataframe wider, so that each spectrum has only one row
Results <- 
  Results %>% 
  pivot_wider(
    id_cols = spectrum_name, 
    names_from = Name, 
    values_from = c(Results_shiny:Dev_toolbox),
    names_sep = "."
    )

##adding a placeholder for the spectrum picture, with the filepath
###filepath
pdf_file_path <- function(spectrum_name) {
  paste0("<img src='Results_ShinyApp/",
    {{ spectrum_name }}, 
    "/",
    {{ spectrum_name }}, 
    "_1_Radiometrie_2022-10-20.png' style=\'height:80px;\'>"
    )
}
###splicing the dataframes together
Results <- cbind(Results[,1], as_tibble_col(
  pdf_file_path(spectra$spectrum_name), column_name = "Picture"), Results[,-1])

```
</details> 

The next chunk prepares the table output in a flexible way:

<details><summary> Setting the Table up </summary> 
```{r}
#setting the table up -------------------

##names for the merging
merging_names <- spectra$excerpt[[1]]$Name
merging_names2 <- spectra$excerpt[[1]]$Name %>% str_replace("\\(", "<br>\\(")

##column names for renaming
col_names <- paste0("Results_shiny.", merging_names)
##creating a list with one entry per variable, named after the column name (to 
##be renamed later)
renaming <- rbind(merging_names2)
names(renaming) <- col_names
renaming <- renaming %>% as.list()
renaming <- map(renaming, md)

#creating a list with cells not to format by decimals
number_fmt_col <- 
  Results %>% select(!starts_with("Dev") & !Picture & !spectrum_name) %>% 
  names()

##function that does the merging
merging <- function(data, Name, condition = "difference") {
  if(condition == "difference"){
  data %>% cols_merge(columns = ends_with(Name, ignore.case = FALSE),
             pattern = "<div style='color:lightgrey'>shiny:</div>{1}<div
             style='color:lightgrey'>luox:</div>{4}<div style='color:lightgrey'>
             toolbox:</div>{5}")
  }
  else {
      data %>% cols_merge(columns = ends_with(Name, ignore.case = FALSE),
             pattern = "<div style='color:lightgrey'>shiny:</div>{1}<div
             style='color:lightgrey'>luox:</div>{2}<div style='color:lightgrey'>
             toolbox:</div>{3}")
  }
}

#creating a function for the gt table
comparison_table <- function(tt_text, fn_text, condition) {
gtobj <- Results %>% 
  gt(rowname_col = c("spectrum_name")) %>% 
  tab_header(title = md(paste0("**Validation Results: ",tt_text , "**"))) %>% 
  tab_footnote(footnote = fn_text)

for(i in seq_along(merging_names)) {
  gtobj <- gtobj %>% merging(merging_names[i], condition = condition)
}

gtobj <- gtobj %>%   
  fmt_markdown(columns = everything()) %>%
  fmt_number(
    columns = all_of(number_fmt_col),
    decimals = 3,
    sep_mark = "",
    pattern = "{x}<br>"
    ) %>%
  cols_align(align = "center") %>% 
  cols_label(.list = renaming) %>%
  sub_missing(missing_text = md("---<br>")) %>% 
  cols_width(
    Picture ~px(150),
    everything() ~ px(80)
  ) %>% 
  opt_align_table_header(align = "left") %>% 
  tab_options(table.font.size = "9px")

gtobj

}

```
</details> 

## Results

This section shows the validation results in two tables. The first table shows the results for the `Shiny` App per spectrum and parameter alongside the relative difference of the respective results from the `luox` app and the `CIE Toolbox`. The second table shows all results per spectrum and parameter. Note that the `Shiny` App does not provide a `CRI [Ra]` for the artificial `EE_Spektrum` and the `LED_4000K_2`, because they exceed the CIE limits for calculation.

::: panel-tabset
### Relative Differences

<details><summary> Creating the gt Table </summary> 
```{r}
#creating the gtable -------------------

#text for the subtitle
fn_text <- 
  md(paste0(
    "The first number in every cell shows the Result from the *Shiny* app, ",
    "<br>the second number the **relative** difference of the respective ",
    "result from the *luox* app, <br>whereas the third number shows the same ",
    "for the result from the *CIE S026 Toolbox*. <br><a style='color:green'>",
    "green</a> values indicate a zero difference, <a style='color:red'>red</a>",
    " a negative difference, and <a style='color:blue'>blue</a> a positive ",
    "difference. <br>All *Shiny* values are rounded to three decimals. <br>", 
    "Missing values or pairwise comparisons are indicated by a ---."))

tt_text <- "Relative Differences"
```
</details> 
```{r}
comparison_table(tt_text, fn_text, condition = "difference")
```

### Results for all sources

<details><summary> Creating the gt Table </summary> 
```{r}

#creating the gtable -------------------

#text for the subtitle
fn_text <- md("The first number in every cell shows the Result from the *Shiny* 
              app, <br>the second number the result from the *luox* app, <br>
              whereas the third number shows the result from the *CIE S026 
              Toolbox*. <br>All values are rounded to three decimals. <br>
              Missing values are indicated by a ---.")

tt_text <- "All Results"
```
</details> 
```{r}
comparison_table(tt_text, fn_text, condition = "")

```
:::

A quick overview of the previous table show that the `Shiny` app produces results that are either identical, or at least very similar to the `luox` app or the `CIE Toolbox`. The next two sections will provide a more concise overview of how those sources compare.

<details><summary> Preparing Data </summary> 
```{r}
#extract the relative difference differentiated by spectrum and variable
Discussion <- 
  spectra %>% 
  select(spectrum_name, excerpt) %>% 
  unnest(excerpt) %>% 
  rowwise() %>% 
  mutate(Dev_luox = 1 - Results_luox/Results_shiny,
         Dev_toolbox = 1 - Results_toolbox/Results_shiny
         ) %>% 
  ungroup()

#throwing the units out for visualization
Discussion <- 
  Discussion %>% mutate(Name = str_replace(Name, "\\(mW ⋅ m⁻²\\)", ""),
                        Name = str_replace(Name, "\\(lx\\)", ""))

```
</details> 

### Pairwise comparison to the `luox` app

<details><summary> luox data </summary> 
```{r}
#creating a subframe for the luox-data, filtered by removing all non-comparisons
Data <- Discussion %>% dplyr::filter(!is.na(Dev_luox))
#number of comparisons made
n <- Data %>% count() %>% pull(1)
#number of comparisons split by difference
n2 <- Data %>% group_by(Dev_luox == 0, Dev_luox > 0, Dev_luox < 0) %>% 
  count() %>% pull(n)
n2[3] <- ifelse(is.na(n2[3]), "none",  n2[3])
#median difference when disregarding sign
n3 <- Data %>% filter(Dev_luox != 0) %>% 
  dplyr::summarise(median = median(abs(Dev_luox))) %>% pull(1)
#maximum difference
n4 <- Data %>% pull(Dev_luox) %>% abs() %>% max()
#where did this difference occur
n4_2 <- Data$Name[abs(Data$Dev_luox) == n4]

```
</details> 

Of `r n` comparisons, `r n2[3]` were *identical*, `r n2[1]` were *smaller*, and `r n2[2]` were *larger*, using the `luox` results as a basis. The *median* relative difference was `r vec_fmt_scientific(n3)` (disregarding sign), i.e., `r vec_fmt_number(n3*100, n_sigfig = 2)`%. The *highest* relative difference (again disregarding sign) was `r  vec_fmt_scientific(n4)` or `r vec_fmt_number(n4*100, n_sigfig = 2)`%, which occured for ``r n4_2``. The following figure provides a histogram of all relative differences (excluding zero difference), colored by variable.

::: panel-tabset
#### Overview

<details><summary> histogram </summary> 
```{r, message = FALSE, warning = FALSE, fig.width = 8}
#make a histogram of the values and calculate relevant values

breaks <- c(10^-15, 10^-10, 10^-5, 1)

Base_Plot <- 
  Data %>% filter(Dev_luox !=0) %>% 
  ggplot(aes(x=abs(Dev_luox))) +
  geom_histogram(aes(fill = Name))+
  xlab("relative difference (irregarding sign)")+
  expand_limits(x= 1)
```
</details> 
```{r, message = FALSE, warning = FALSE, fig.width = 8}
Base_Plot +
  scale_x_log10(breaks = breaks, labels = c(vec_fmt_number(breaks*100, 
                                                           n_sigfig = 1, 
                                                           pattern = "{x}%"))) +
  ylab("no. of spectra")
```

#### Split by Variable

```{r, message = FALSE, warning = FALSE, fig.width = 8}
Base_Plot +
  scale_x_log10(breaks = breaks)+
  ylab("no. of spectra")+
  facet_wrap("Name")
```

#### Split by Spectrum

```{r, message = FALSE, warning = FALSE, fig.width = 8}

Base_Plot +
  scale_x_log10(breaks = breaks)+
  ylab("no. of variables")+
  facet_wrap("spectrum_name")

```
:::

### Pairwise comparison to the `CIE Toolbox`

<details><summary> CIE prep </summary> 
```{r}
#creating a subframe for the luox-data, filtered by removing all non-comparisons
Data <- Discussion %>% dplyr::filter(!is.na(Dev_toolbox))
#number of comparisons made
n <- Data %>% count() %>% pull(1)
#number of comparisons split by difference
n2 <- Data %>% group_by(Dev_toolbox == 0, Dev_toolbox > 0, Dev_toolbox < 0) %>% 
  count() %>% pull(n)
n2[3] <- ifelse(is.na(n2[3]), "none",  n2[3])
#median difference when disregarding sign
n3 <- Data %>% filter(Dev_toolbox != 0) %>% 
  dplyr::summarise(median = median(abs(Dev_toolbox))) %>% pull(1)
#maximum difference
n4 <- Data %>% pull(Dev_toolbox) %>% abs() %>% max()
#where did this difference occur
n4_2 <- Data$Name[abs(Data$Dev_toolbox) == n4]

```
</details> 

Of `r n` comparisons, `r n2[3]` were *identical*, `r n2[1]` were *smaller*, and `r n2[2]` were *larger*, using the `CIE Toolbox` results as a basis. The *median* relative difference was `r vec_fmt_scientific(n3)` (disregarding sign), i.e., `r vec_fmt_number(n3*100, n_sigfig = 2)`%. The *highest* relative difference (again disregarding sign) was `r  vec_fmt_scientific(n4)` or `r vec_fmt_number(n4*100, n_sigfig = 2)`%, which occured for ``r n4_2``. The following figure provides a histogram of all relative differences (excluding zero difference), colored by variable.

::: panel-tabset
#### Overview

<details><summary> Histogram </summary> 
```{r,  message = FALSE, warning = FALSE, fig.width = 8}
#make a histogram of the values and calculate relevant values

Base_Plot <- 
  Data %>% filter(Dev_toolbox !=0) %>% 
  ggplot(aes(x=abs(Dev_toolbox))) +
  geom_histogram(aes(fill = Name))+
  xlab("relative difference (irregarding sign)")+
  expand_limits(x= 1)
```
</details> 

```{r, message = FALSE, warning = FALSE, fig.width = 8}
Base_Plot +
  scale_x_log10(breaks = breaks, labels = c(vec_fmt_number(breaks*100, 
                                                           n_sigfig = 1, 
                                                           pattern = "{x}%")))+
  ylab("no. of spectra")
```

#### Split by Variable

```{r, message = FALSE, warning = FALSE, fig.width = 8}

Base_Plot +
  scale_x_log10(breaks = breaks)+
  ylab("no. of spectra")+
    facet_wrap("Name")
```

#### Split by Spectrum

```{r, message = FALSE, warning = FALSE, fig.width = 8}

Base_Plot +
  scale_x_log10(breaks = breaks)+
  ylab("no. of variables")+
  facet_wrap("spectrum_name")

```
:::

## Conclusion {#sec-conclusion}

Overall, the agreement between the different sources is very high and differences only occur several decimals back. For most if not all of those cases, rounding errors seem a plausible explanation. In an older version of the `Shiny` app, negative input values for irradiance were taken at *face value*, i.e., they actually reduced variable values that require summation. With that state, many more comparisons showed a zero difference for the `luox` app, which seems to indicate that the *luox app* also takes negative values at *face value*. However, the median relative difference in that older state was about double that compared to the current state for both the `luox` app and the `CIE Toolbox` sources. Since both the overall error is reduced by the current state of the `Shiny` app and it is sensible to restrict input values to zero or positive numbers, this method will be used in the public release.

In summary, the `Shiny` app offers a sufficiently accurate calculation of ⍺-opic values, especially given its focus on education. It is of note, however, that the age corrected values could no be validated against the other two sources, since they don´t provide similar functionality.
